Monitoring forest health using hyperspectral imagery: Does feature selection improve the performance of machine-learning techniques?

### 2021-06-07 docs/00-manuscripts/ieee/ieee-manuscript.tex
edited response to reviewer 1
fixed affiliations according to corporate style guides

\author{Patrick~Schratz,
	Jannes~Muenchow,
	Eugenia~Iturritxa,
	José~Cortés,
	Bernd~Bischl,
	and Alexander~Brenning
	\thanks{P.Schratz, J.Muenchow, J.Cortés and A.Brenning are with the Department
		of Geography, GIScience group, Friedrich-Schiller-University of Jena, Germany.}% <-this % stops a space
	\thanks{B.Bischl is head of the computational statistics group at the Department of Statistics, Ludwig-Maximilian-University Munich.}% <-this % stops a space
	\thanks{E.Iturritxa is with NEIKER Tecnalia, Vitoria-Gasteiz, Arab, Spain.}% <-this % stops a space
	%\thanks{Manuscript received April 19, 2005; revised August 26, 2015.}
}

@@

Vegetation indices have shown the potential to provide valuable information when analyzing forest health \cite{jiang2014, adamczyk2015}.
Most vegetation indices were developed with the aim of being sensitive to changes of specific wavelength regions, serving as a proxy for underlying plant processes.
However, indices developed/applied for different purposes than the one to be analyzed (i.e. defoliation at pine trees) may help to explain complex underlying relationships which are not obvious on the first look.
This emphasizes the need to extract as much information as possible from the available input data to generate promising features which can help to understand the modeled relationship \cite{thenkabail2018}.
A less known index type which can be derived from spectral information is the \ac{NRI}.
In contrast to most vegetation indices, \ac{NRI}s do not use an expert-based formula following environmental heuristics but instead makes use of a data-driven feature engineering approach by combining (all possible) combinations of spectral bands \cite{thenkabail2000}.
Especially when working with hyperspectral data, thousands of \ac{NRI} features can be derived this way.

@@

	is a Data Scientist working in Zurich, CH.

@@

	He is a PhD candidate at the GIScience group, Department of Geography at University of Jena where he acts as a researcher in the environmental modeling field.

@@

	He has earned his PhD at the Friedrich-Alexander University Erlangen-Nuremberg (Germany) in 2013.

@@

	received the Ecology and Ph.D degree in Plant Protection from the Basque country University in 2001.

@@

	Her research includes the distribution of diseases, analysis of predisposing factors for them, genetic and phenotypic studies of populations and their epidemiology.

@@

	Currently he is a PhD student at Friedrich-Schiller-University, Jena, Germany.

@@

	He is a professor for "Statistical Learning and Data Science" at the Department of Statistics at the Ludwig-Maximilians-University Munich and a director of the "Munich Center for Machine Learning".

@@

	He served as an assistant professor and tenured associate professor in geomatics at the University from 2007 until 2015, when he was appointed as a full professor in geographic information science at Friedrich Schiller University Jena, Germany.

>>>>

\author{Patrick~Schratz,
	Jannes~Muenchow,
	Eugenia~Iturritxa,
	José~Cortés,
	Bernd~Bischl,
	and Alexander~Brenning
	\thanks{P.Schratz, J.Muenchow, J.Cortés and A.Brenning are with the Department
		of Geography, geographic information science group, Friedrich Schiller University Jena, Germany.}% <-this % stops a space
	\thanks{B.Bischl is head of the computational statistics group at the Department of Statistics, Ludwig-Maximilian-Universität in Munich, Germany.}% <-this % stops a space
	\thanks{E.Iturritxa is with NEIKER Tecnalia, Vitoria-Gasteiz, Spain.}% <-this % stops a space
	%\thanks{Manuscript received April 19, 2005; revised August 26, 2015.}
}

@@

Vegetation indices have shown the potential to provide valuable information when analyzing forest health \cite{jiang2014, adamczyk2015}.
Most vegetation indices were developed with the aim of being sensitive to changes of specific wavelength regions, serving as a proxy for underlying plant processes.
However, indices developed/applied for different purposes than the one to be analyzed (i.e. defoliation at pine trees) may help to explain complex underlying relationships which are not obvious at first.
This emphasizes the need to extract as much information as possible from the available input data to generate promising features which can help to understand the modeled relationship \cite{thenkabail2018}.
A less known index type which can be derived from spectral information is the \ac{NRI}.
In contrast to most vegetation indices, \ac{NRI}s do not use an expert-based formula following environmental heuristics but instead makes use of a data-driven feature engineering approach by combining (all possible) combinations of spectral bands \cite{thenkabail2000}.
Especially when working with hyperspectral data, thousands of \ac{NRI} features can be derived this way.

@@

	is a Data Scientist working in Zurich, Switzerland.

@@

	He is a PhD candidate in geographic information science in the Department of Geography at Friedrich Schiller University Jena where he conducts research in environmental modeling.

@@

	He received his PhD from Friedrich-Alexander University Erlangen-Nürnberg (Germany) in 2013.

@@

	received the Ecology and Ph.D degree in Plant Protection from the University of the Basque Country in 2001.

@@

	Her research the NEIKER research center includes the distribution of diseases, analysis of predisposing factors for them, genetic and phenotypic studies of populations and their epidemiology.

@@

	Currently he is a PhD student at Friedrich Schiller University, Jena, Germany.

@@

	He is a professor of Statistical Learning and Data Science in the Department of Statistics at Ludwig-Maximilians-Universität in Munich, Germany, and a director of the Munich Center for Machine Learning.

@@

	He served as an assistant professor and tenured associate professor in geomatics at the University of Waterloo, Ontario, Canada from 2007 until 2015, when he was appointed as a full professor in geographic information science at Friedrich Schiller University Jena, Germany.
